Subglacial and subaerial fluvial sediment transport capacity respond differently to water discharge variations

### 2024-08-05 Delaneyetal2024.tex
changes

%% 2-column papers and preprints
\documentclass[esurf, manuscript]{copernicus}

@@

\begin{abstract}
  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom.
  This quantity varies with water velocity and channel width.
  In subaerial channels,  sediment transport capacity can co-vary with water discharge because its variations are accommodated by changes in flow depth, width, and water velocity.
  By contrast, subglacial channels evolve at a slower rate than most water discharge variations, causing a variable relationship between channel shape, water velocity, and discharge.
  Using  sediment transport formulations and numerical models applied to two observed hydrographs, we evaluate the differences in sediment transport capacity variability between subaerial and subglacial channels.
  The numerical experiments confirm that subglacial sediment transport is highly non-linear with respect to water discharge.
  Yet, subglacial sediment transport capacity can vary less if water discharge variations are closer to equilibrium with subglacial channel size.
  In a dedicated numerical experiment, we show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
  The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
  This implies that hysteresis in channel size causes a variable relationship between sediment and water discharge in a transport-limited subglacial system.
  Lastly, we present formulations that demonstrate the impact of different assumptions of channel behavior on sediment transport capacity.
  The implications are discussed for sediment discharge from glaciers with different hydro-climatic forcings.
  We also evaluate the conditions under which subglacial sediment transport capacity could co-vary with water discharge.
  These findings can also improve the interpretation of sediment discharge records in glacierized catchments.
\end{abstract}

@@

The manuscript's first objective is to establish if water discharge covaries with subglacial sediment transport capacity.
In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during an discharge event, due to evolving channel size.
In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.
Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected in most systems.
This incoherent relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
Water discharge and sediment transport capacity covariance could be possible when water discharge varies at a slower rate than subglacial channel size.
In natural environments, this appears to rarely be the case.

>>>>

%% 2-column papers and preprints
\documentclass[tc, manuscript]{copernicus}

@@

\begin{abstract}
  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom.
  This quantity varies with water velocity and channel width.
  In subaerial channels,  sediment transport capacity can co-vary with water discharge because its variations are accommodated by changes in flow depth, width, and water velocity.
  By contrast, subglacial channels evolve at a slower rate than most water discharge variations, causing a variable relationship between channel shape, water velocity, and discharge.
  Using  sediment transport formulations and numerical models applied to two observed hydrographs, we evaluate the conditions where sediment transport capacity varies with water discharge in subglacial channels.
  Numerical experiments show that the changing channel size results in sediment transport capacity peaking before the maximum water discharge.
  This hysteresis in channel size causes a highly variable relationship between sediment and water discharge in a transport-limited subglacial system.
  The results also indicate that high subglacial sediment transport capacities can occur across a wide range of water discharges.
  We  then evaluate the conditions where subglacial sediment transport capacity is more variable than its subaerial counterpart.
  The numerical experiments of real hydrographs confirm that subglacial sediment transport is highly non-linear with respect to water discharge, creating more variability in sediment transport capacity.
  Yet,  formulations of subglacial sediment transport capacity show that it can covary with water discharge variations when subglacial channel size is in equilibrium. 
  The implications of these findings are discussed in the context of sediment discharge from glaciers with different hydro-climatic forcings.
  We also discuss the impact of different assumptions of channel behavior on sediment transport capacity.
  These findings can improve the interpretation of sediment discharge records in glacierized catchments.
\end{abstract}

@@

The manuscript's first objective is to establish if water discharge covaries with subglacial sediment transport capacity.
In subglacial channels, the timing of peak sediment discharge capacity occurs before peak water discharge during an discharge event, due to evolving channel size.
In subaerial channels, the timing of peak sediment discharge capacity and water discharge coincide.
Results here suggest that, even in a transport-limited subglacial system, an incoherent relationship between water and subglacial sediment discharge could be expected in most channels.
This incoherent relationship presents a challenge in linking hydro-climatic conditions or events to sediment export from glaciers.
Water discharge and sediment transport capacity covariance could be possible when water discharge varies at a slower rate than subglacial channel size.
In natural environments, this appears to rarely be the case.
